```{r data generation, echo = FALSE, results = "hide"}
names = read.csv("names2.txt", header = FALSE, sep = ",")

item = c("type of crust","type of crust","size","sauce")
items = c("types of crust","types of crust","sizes","sauces")


while(T){
  ns = sample(2:5,2,replace=T)
  ns[3] = ns[2]-1
  if(prod(ns) < 21){
    break
  }
}

na = sample(names,1)
i = 1:length(item)
it = item[i]
its = items[i]
```

Question
========

`r na` is buying a pizza. `r na` still has to decide on a `r it[1]` and 2 toppings. The shop carries `r ns[1]` `r its` and `r ns[2]` toppings. How many different combinations is `r na` choosing between?


Solution
========

Make a tree. Notice that the second topping has one fewer choices. Also, 

```{r tree, echo=F}
par(mar=c(0,0,0,0))
plot(1,1,type ="n",axes=F,ann=F,xlim=c(0,1),ylim=c(0,1.1))
a = ns[1]
b = prod(ns[1:2])
c = prod(ns)

as = (2*(1:a)-1)/(a*2)
bs = (2*(1:b)-1)/(b*2)
cs = (2*(1:c)-1)/(c*2)

# cs = 1:prod(ns)/(prod(ns)+1)
# bs = 1:prod(ns[1:2])/(prod(ns[1:2])+1)
# as = 1:ns[1]/(prod(ns[1])+1)
# points(0,0.5)
# points(rep(0.33,length(as)),as)
# points(rep(0.66,length(bs)),bs)
# points(rep(1,length(cs)),cs)
for(i in 1:ns[1]){
  lines(c(0,0.33),c(0.5,as[i]))
  for(j in 1:ns[2]){
    lines(c(0.33,0.66),c(as[i],bs[(i-1)*ns[2]+j]))
    for(k in 1:ns[3]){
      lines(c(0.66,1),c(bs[(i-1)*ns[2]+j],cs[((i-1)*ns[2]+j-1)*ns[3]+k]))
    }
  }
}
text(0.1666,1.05,it)
text(0.33+0.1666,1.05,"first topping")
text(0.666+0.1666,1.05,"second topping")
```

Count the leaves (the nodes at the far right). In this case there are `r c` leaves.

$$\fbox{There are `r c` combinations possible.} $$


Meta-information
============
extype: string
exsolution: yup
exname: reading hist
extol: 0.01
