```{r data generation, echo = FALSE, results = "hide"}
n= sample(3:4,1)
p = round(runif(1,0.20,0.80),2)

mu = sum(n*p)
sig = sqrt(n*p*(1-p))

u = runif(n,0,1)
h = u>p
x = sum(h)
```

Question
========

Determine the probabilities when adding up $`r n`$ Bernoulli trials if each trial has chance $`r p`$. Each trial could be thought of as a spin of the spinner below.

<!-- Also, measure the sums of $`r n`$ spins 10 times, and report the results as a histogram.  -->


\setkeys{Gin}{width=.3\textwidth}

\tiny

```{r wheel, echo=F,fig.height=2, fig.width=2, fig.cap=''}
par(mar=c(0,0,0,0))
plot(1,1,xlim=c(-1,1),ylim=c(-1,1),ann=F,axes=F,type='n')
thetas = 1:1000 / 1000 * 2 * pi
xs = cos(thetas)
ys = sin(thetas)
polygon(xs,ys)
q = 1-p

lines(c(0,cos(q*2*pi)),c(0,sin(q*2*pi)))
lines(c(0,1),c(0,0),lwd=2)

text(0.8*cos(mean(c(0,q))*2*pi),0.8*sin(mean(c(0,q))*2*pi),0)
text(0.8*cos(mean(c(q,1))*2*pi),0.8*sin(mean(c(q,1))*2*pi),1)

arrows(-0.6,0,0.6,0,lwd=4,length=0.1)
```

\normalsize

Then, use $\mu = \sum x\cdot \text{Pr}(x)$ to find the mean and $\sigma = \sqrt{\sum (x-\mu)^2 \cdot \text{Pr}(x)}$ to determine the standard deviation.

Then, show $\mu = np$ and $\sigma = \sqrt{npq}$. (Remember these formulas only work for binomial distributions.)

Solution
========

\small

You could make a tree.


\setkeys{Gin}{width=.8\textwidth}

\tiny

```{r tree, echo=F,fig.width=2*6,fig.height=6,fig.cap=''}
par(mar=c(0,0,0,0))
plot(2,1,xlim=c(0,2),ylim=c(0,1),ann=F,axes=F,type='n')
x = numeric()
y = numeric()
for(i in 1:(n)){
  mm = 2^(i-1)
  for(j in 1:mm){
    x = (i-1)/(n)
    y = (2*j-1)/2/mm
    x2 = (i)/(n)
    x3 = x2
    j2 = 2*j-1
    j3 = 2*j
    y2 = (2*j2-1)/2/2^i
    y3 = (2*j3-1)/2/2^i
    lines(c(x,x2),c(y,y2),lty=2)
    lines(c(x,x3),c(y,y3),lty=2)
    points(x,y)
    points(x2,y2)
    points(x3,y3)
    text(mean(c(x,x2)),mean(c(y,y2)),p)
    text(mean(c(x,x3)),mean(c(y,y3)),1-p)
  }
}

number2binary = function(number, noBits) {
       binary_vector = rev(as.numeric(intToBits(number)))
       if(missing(noBits)) {
          return(binary_vector)
       } else {
          binary_vector[-(1:(length(binary_vector) - noBits))]
       }
}

for(i in 0:((2^n)-1)){
  x4 = number2binary(i,n)
  s4 = ''
  for(j in x4){
    s4 = paste(s4,j)
  }
  text(1.25, 1-(2*i+1)/2/2^n, paste("Pr(",s4,") = ",round(p^(sum(x4))*(1-p)^(n-sum(x4)),4),sep=''))
  lines(c(1.4,1.6),c(1-(2*i+1)/2/2^n,1-(sum(x4)+1)/(n+2)))
}

for(i in 0:n){
  text(1.75,1-(i+1)/(n+2),paste("Pr(",i,")=",signif(dbinom(i,n,p),3),sep=""))
}
```

\small


You could also just use the binomial formula.
$$\text{Pr}(x) = {_nC_x}(p)^x(1-p)^{n-x}$$

We want `r n+1` probabilities, letting $x$ vary from 0 to `r n`. For each probability, $n=`r n`$ and $p=`r p`$. A table is useful. 

```{r tabs, echo=F}
ss = ''
for(x in 0:n){
  ss = paste(ss,"|",x,"|$({_",n,"C_",x,"})(",p,")^{",x,"}(1-",p,")^{",n,"-",x,"}$ |",signif(dbinom(x,n,p),3),"|\n",sep="")
}
```

| $x$ | ${_nC_x}p^x(1-p)^{n-x}$ | $\text{Pr}(x)$ |
|:---:|:---:|:---:|
`r ss`

```{r meansd, echo=F}
x = 0:n
P = signif(dbinom(x,n,p),3)
mu = signif(sum(P*x),4)
sig = signif(sqrt(sum(P*(x-mu)^2)),2)
c3 = signif(x*P,3)
c4 = signif(x-mu,3)
c5 = signif((x-mu)^2,3)
c6 = signif(P*(x-mu)^2,3)
s2 = ''
for(i in 1:(n+1)){
    s2 = paste(s2,'|',x[i],'|',P[i],'|',c3[i],'|',c4[i],'|',c5[i],'|',c6[i],'|')
    if(i < n+1){
        s2 = paste(s2,'\n')
    }
}
```


| $x$ | ${Pr}(x)$ | $x \cdot {Pr}(x)$ |$x-\mu$| $(x-\mu)^2$ | $(x-\mu)^2 \cdot {Pr}(x)$ | 
|:---:|:---:|:---:|:---:|:---:|:---:|
`r s2`
|=========|=========|=========|=========|=========|=========|
|     |     | $\sum x \cdot {Pr}(x) = `r mu`$ | |  | $\sum (x_i-\mu)^2 \cdot {Pr}(x) = `r signif(sum(c6),3)`$   |
|     |     | $\mu = `r mu`$ | |   | $\sigma = \sqrt{\sum (x_i-\mu)^2 \cdot {Pr}(x)} = `r sig`$   |

Then we confirm the binomial mean and binomial SD formulas work.
$$\mu = np = (`r n`)(`r p`) = `r n*p` $$
$$\sigma = \sqrt{npq} = \sqrt{(`r n`)(`r p`)(`r q`)} = \sqrt{`r signif(n*p*q,3)`} = `r signif(sqrt(n*p*q),3)` $$

\normalsize


Meta-information
============
extype: string
exsolution: yo
exname: marbles
extol: 0.01
