---
output:
  pdf_document: default
  html_document: default
---
```{r data generation, echo = FALSE, results = "hide", fig.show="hide"}
## DATA GENERATION

types = c("bimodal","symmetric mound","skew left","skew right","uniform")
shapes1 = c(0.5, 20,   3,   1, 1)
shapes2 = c(0.5, 20,   1,   3, 1)

rrr = sample(1:5,1)
type = types[rrr]
shape1 = shapes1[rrr]
shape2 = shapes2[rrr]

n = sample(6:12,1)
offset = sample(1:9*10,1)+0.001
rng = sample(c(3,5,10,20,40,100),1)
data = rbeta(n,shape1,shape2)*rng*0.99 + offset
index = 1:n
x = sort(round(data,3))

is = sample(1:n,2)
ps = signif(is/n,3)
```

Question
========

A continuous random variable $X$ was measured `r n` times. The sorted data are shown below, along with each datum's index.

```{r make table, echo=FALSE}
st = ''
for(i in 1:n){
  st = paste(st, sprintf("|%d|%.3f|\n",i,x[i]))
}
```
| $i$ | $x$ |
|:---:|:-----:|
`r st`

Answerlist
--------
* Determine the percentile rank of the value `r x[is[1]]`. In other words, determine what percent of data are less than or equal to `r x[is[1]]`.
* Determine the datum corresponding to a percentile rank of `r ps[2]`. In other words, determine $x$ such that `r ps[2]*100`% of the data are less than or equal to $x$. 

Solution
========

Let $x$ represent a datum of interest. Let $i$ represent that datum's index. Let $\ell$ represent that datum's percentile. Let $n$ represent the sample size (number of measurements). In general,
$$\ell = \frac{i}{n} $$


Answerlist
--------
* We are given $x = `r x[is[1]]`$. This means $i=`r is[1]`$. We know $n=`r n`$. Determine the percentile $\ell$.
$$\begin{aligned}
\ell &= \frac{`r is[1]`}{`r n`} \\ \\
\ell &= `r ps[1]`
\end{aligned}$$
So, the answer is `r ps[1]`, or `r ps[1]*100`%.
* We are given $\ell = `r ps[2]`$. We can use algebra to solve for $i$.
$$\ell = \frac{i}{n}$$
Multiply both sides by $n$.
$$n \cdot (\ell) = n \cdot \left(\frac{i}{n} \right)$$
Simplify both sides.
$$ n \ell = i $$
To make me happy, switch the sides.
$$i = n \ell $$
Now, we can evaluate $i$.
$$i = (`r n`)(`r ps[2]`) $$
$$i = `r is[2]`$$
Determine the $x$ associated with $i = `r is[2]`$.
$$x = `r x[is[2]]`$$


Meta-information
============
extype: string
exsolution: yup
exname: reading hist
extol: 0.01
