```{r data generation, echo = FALSE, results = "hide"}
## DATA GENERATION
library(latex2exp)
options(scipen=999)
alpha = sample(c(0.2,0.1,0.05,0.04,0.02,0.01),1)
zstar = abs(qnorm(alpha/2))
while(T){
    p = round(runif(1,0.05,0.35),3)
    n = sample(1:10,1) * 10^sample(0:3,1)
    se = signif(sqrt(p*(1-p)/n),3)
    if (n*p>10 & n*(1-p)>10 & sqrt(n*p*(1-p))>7)
    {
        break
    }
}
z = round(runif(1,zstar-0.3,zstar+0.3),2) * sample(c(-1,1),1)
phat = signif(p+z*se,3)
ns = round(phat*n)
phat = round(ns/n,3)
z = round((phat-p)/sqrt(p*(1-p)/n),2)
Pr = 2*round(pnorm(-abs(z)),4)
wrd2 = sample(c('not', 'not equal to', 'different than'),1)
if (Pr<alpha)
{
    wrd3 = 'reject'
    sym = '<'
    ans = 'We conclude the candy maker is wrong.'
} else {
    wrd3 = "don't reject"
    sym = '>'
    ans = "We conclude the candy maker could be correct."
}

```

Question
========

A candy maker claims `r p*100`% of the candies are purple. We decide to test this claim with a `r alpha` significance level. We collect a random sample of `r n` candies and `r ns` of them are purple.

Answerlist
--------
* What kind of hypothesis test is appropriate?
* State the hypotheses
* Determine the $p$-value.
* Decide whether we reject or retain the null hypothesis.
* Make a conclusion.


Solution
========

We should use a two-tail proportion test.

State the hypotheses.
$$H_0 ~~\text{ claims } ~~ p = `r p` $$
$$H_\text{A} ~~\text{ claims } ~~ p \ne `r p`$$
Determine the standard error.
$$\sigma_{\hat{p}} = \sqrt{\frac{p_0(1-p_0)}{n}} = \sqrt{\frac{`r p`(1-`r p`)}{`r n`}} = `r se` $$
Determine the sample proportion.
$$\hat{p} = `r phat` $$
Determine a $z$ score. For simplicity, we ignore the continuity correction.
$$z = \frac{\hat{p}-p_0}{\sigma_{\hat{p}}} = \frac{`r phat`-`r p`}{`r se`} = `r z` $$
Make a sketch of the null's sampling distribution. The $p$-value is a two-tail area.

```{r p_single_test_twotail_coin, echo = FALSE, results = "hide", fig.height = 3, fig.width = 6, fig.cap = ""}
par(mar=c(5,3,1,1))
zs = seq(-3,3,length=1000)
ys = dnorm(zs)
plot(zs, ys, 'l',xlim = c(-3,3), ylim = c(0,0.4),
     ann=FALSE,xaxt="n",yaxt="n",
     lwd=1,bty='n')
axis(side=1, line=0, at=-3:3, labels = FALSE)
mtext(signif(p+(-3:3)*se,3),side=1,line=0.5,at=-3:3)
mtext(TeX("\\hat{p}="),line=0.5,side=1,at=-3.6)
axis(side=1, line=2,at=-3:3, labels = FALSE)
mtext(-3:3,side=1,line=2.5,at=-3:3)
mtext(c("z="),line=2.5,side=1,at=-3.6)
polygon(c(zs[zs<=-abs(z)], -abs(z), -3),c(ys[zs<=-abs(z)], 0, 0), col=rgb(1,0,0,0.5))
polygon(c(zs[zs>=abs(z)], 3, abs(z)),c(ys[zs>=abs(z)], 0, 0), col=rgb(1,0,0,0.5))
lines(c(-3,3),c(0,0),'l',lwd=1)
```
\
To determine that two-tail area, we use the z table.
$$\begin{aligned}
p\text{-value} &= P(|Z|>`r abs(z)`) \\
               &= 2\cdot P(Z<`r -abs(z)`) \\
               &= `r Pr` 
\end{aligned}$$

Compare $p$-value to $\alpha$ (which is `r alpha`).
$$p\text{-value} `r sym` \alpha $$
Make the conclusion: we `r wrd3` the null hypothesis.

`r ans`


Answerlist
--------
* Two-tail proportion test
* Hypotheses: $H_0 \text{ claims } p = `r p`$  and $H_\text{A} \text{ claims } p \ne `r p`$.
* The $p$-value is `r Pr`
* We `r wrd3` the null hypothesis.
* `r ans`


Meta-information
============
extype: string
exsolution: `r Pr`
exname: p hyp test twotail
extol: 0.01
