---
output:
  pdf_document: default
  html_document: default
---
```{r data generation, echo = FALSE, results = "hide"}
zs = (-3*5):(3*5)/5
ys = 0:8*0.05
z = sample(1:(2*5)/5,1)
mu = sample(10:60,1)
si = sample(1:round(mu/3),1)
```

Question
========

Let $X$ be a normal random variable with mean $\mu = `r mu`$ and standard deviation $\sigma = `r si`$.
$$X \sim {\cal N}(`r mu`,\,`r si`) $$
The figure below shows the density of random variable $X$. Each grid square represents 1% of probability.

```{r gridnorm,echo=F,fig.height=4,fig.width=10,fig.cap="",fig.path=paste("ug",sample(1:99999,1),sep="")}
plot(0,0,type="n",ann=F,axes=F,xlim=c(-3,3),ylim=c(0,0.4))
abline(h = ys,lty=2)
abline(v=zs,lty=2)
lines((-300:300)/100,dnorm((-300:300)/100))
axis(1,at=(-3):(3),labels=-3:3*si+mu)
mtext("X",side=1,line = 3)
lines(c(-3,3),c(0,0))
```
\

Answerlist
--------
* Estimate $a$ such that $P(|X-\mu|>a) = `r round(2-2*pnorm(z),2)`$ by shading and counting.
* Determine $z$ such that $P(|X-\mu|>a) = `r round(2-2*pnorm(z),2)`$ by using the $z$-table.


Solution
========

Answerlist
--------

* The shaded regions are shown below. \newline
```{r gridnormsol,echo=F,fig.height=4,fig.width=10,fig.cap="",fig.path=paste("ug",sample(1:99999,1),sep="")}
plot(0,0,type="n",ann=F,axes=F,xlim=c(-3,3),ylim=c(0,0.4))
abline(h = ys,lty=2)
abline(v=zs,lty=2)
xs = (-300:300)/100
dn = dnorm(xs)
lines(xs,dn)
lines(c(-3,3),c(0,0))
polygon(c(xs[xs<=-z],-z,-3),c(dn[xs<=-z],0,0),col = rgb(0.4,0.4,0.4,0.4))
polygon(c(xs[xs>=z],3,z),c(dn[xs>=z],0,0),col = rgb(0.4,0.4,0.4,0.4))
axis(1,at=(-3):(3),labels=-3:3*si+mu)
mtext("X",side=1,line = 3)
``` 
\
\newline
When you have shaded `r round(2-2*pnorm(z),2)*100` squares, starting at both tails, you should end near $x=`r z*si+mu`$. Really, you want to shade `r round(round(2-2*pnorm(z),2)*100/2,3)` squares starting from the left and also `r round(round(2-2*pnorm(z),2)*100/2,3)` squares starting from the right. This gives $a = \fbox{`r z*si`}$.
* $a=\fbox{`r round(-qnorm(round(2-2*pnorm(z),2)/2),2)*si`}$



Meta-information
============
extype: string
exsolution: yo
exname: marbles
extol: 0.01
