---
output:
  pdf_document: default
  html_document: default
---
```{r data generation, echo = FALSE, results = "hide"}
zs = (-3*5):(3*5)/5
ys = 0:8*0.05

z = sample((-1.4*5):(1.4*5)/5,1)

mu = sample(10:60,1)
si = sample(1:round(mu/3),1)
```

Question
========

Let $X$ be a normal random variable with mean $\mu = `r mu`$ and standard deviation $\sigma = `r si`$.
$$X \sim {\cal N}(`r mu`,\,`r si`) $$
The figure below shows the density of random variable $X$. Each grid square represents 1% of probability.

```{r gridnorm,echo=F,fig.height=4,fig.width=10,fig.cap="",fig.path=paste("ug",sample(1:99999,1),sep="")}
plot(0,0,type="n",ann=F,axes=F,xlim=c(-3,3),ylim=c(0,0.4))
abline(h = ys,lty=2)
abline(v=zs,lty=2)
lines((-300:300)/100,dnorm((-300:300)/100))
axis(1,at=(-3):(3),labels=-3:3*si+mu)
mtext("X",side=1,line = 3)
lines(c(-3,3),c(0,0))
```
\

Answerlist
--------
* Estimate $P(X<`r z*si+mu`)$ by shading and counting.
* Determine $P(Z<`r z*si+mu`)$ by using the $z$-table.


Solution
========

Answerlist
--------

You will want a $z$-score.
$$z = \frac{x-\mu}{\sigma} = `r z`$$

* The shaded region is shown below. \newline
```{r gridnormsol,echo=F,fig.height=4,fig.width=10,fig.cap="",fig.path=paste("ug",sample(1:99999,1),sep="")}
plot(0,0,type="n",ann=F,axes=F,xlim=c(-3,3),ylim=c(0,0.4))
abline(h = ys,lty=2)
abline(v=zs,lty=2)
xs = (-300:300)/100
dn = dnorm(xs)
lines(xs,dn)
lines(c(-3,3),c(0,0))
polygon(c(xs[xs<=z],z,-3),c(dn[xs<=z],0,0),col = rgb(0.4,0.4,0.4,0.4))
axis(1,at=(-3):(3),labels=-3:3*si+mu)
mtext("X",side=1,line = 3)
```
\
\newline
You should count about `r round(pnorm(z),2)*100` shaded squares, giving a probability of about `r round(pnorm(z),2)`.
* The probability is `r round(pnorm(z),4)`. 



Meta-information
============
extype: string
exsolution: yo
exname: marbles
extol: 0.01
