---
header-includes:
   - \usepackage{multicol}
output:
  pdf_document: default
  html_document: default
---

```{r data generation, echo = FALSE, results = "hide"}

makepie = function(p1,p2){ #,name){
  #pdf(name,width=2.5,height=2.5)
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",ann=F,axes=F,xlim=c(-1.1,1.1),ylim=c(-1.1,1.1))
  ps = 0:1000/1000
  the = ps*2*pi
  lines(cos(the)*0.8,sin(the)*0.8)
  if(p1<p2){
    polygon(c(0,cos(the[ps>=p1 & ps<=p2])*0.8,0),c(0,sin(the[ps>=p1 & ps<=p2])*0.8,0),col=rgb(0.4,0.4,0.4,0.4),lty=0)
    pave = (p1+p2)/2
    text(0,0,paste("P=",round((p2-p1),3),sep=""))
    text(cos(p1*2*pi)*0.5,sin(p1*2*pi)*0.5,round(p1,3))
    text(cos(p2*2*pi)*0.5,sin(p2*2*pi)*0.5,round(p2,3))
    lines(c(cos(p1*2*pi)*0.7,cos(p1*2*pi)*0.83),
          c(sin(p1*2*pi)*0.7,sin(p1*2*pi)*0.83),
          lwd = 2)
    lines(c(cos(p2*2*pi)*0.7,cos(p2*2*pi)*0.83),
          c(sin(p2*2*pi)*0.7,sin(p2*2*pi)*0.83),
          lwd = 2)
    text(cos(p1*2*pi)*1.05,sin(p1*2*pi)*1.05,round(qnorm(p1),2))
    text(cos(p2*2*pi)*1.05,sin(p2*2*pi)*1.05,round(qnorm(p2),2))
  }
  if(p1>p2){
    polygon(c(0,cos(the[ps>p1])*0.8,0),c(0,sin(the[ps>p1])*0.8,0),col=rgb(0.4,0.4,0.4,0.4),lty=0)
    polygon(c(0,cos(the[ps<p2])*0.8,0),c(0,sin(the[ps<p2])*0.8,0),col=rgb(0.4,0.4,0.4,0.4),lty=0)
    pave = (p2+1-p1)/2+p1
    text(0,0,paste("P=",round((p2+1-p1),3),sep=""))
    text(cos(p1*2*pi)*0.5,sin(p1*2*pi)*0.5,round(p1,3))
    text(cos(p2*2*pi)*0.5,sin(p2*2*pi)*0.5,round(p2,3))
    lines(c(cos(p1*2*pi)*0.7,cos(p1*2*pi)*0.83),
          c(sin(p1*2*pi)*0.7,sin(p1*2*pi)*0.83),
          lwd = 2)
    lines(c(cos(p2*2*pi)*0.7,cos(p2*2*pi)*0.83),
          c(sin(p2*2*pi)*0.7,sin(p2*2*pi)*0.83),
          lwd = 2)
    text(cos(p1*2*pi)*1.05,sin(p1*2*pi)*1.05,round(qnorm(p1),2))
    text(cos(p2*2*pi)*1.05,sin(p2*2*pi)*1.05,round(qnorm(p2),2))
  }
  
  #dev.off()
}


options(digits = 4)

u2 = round(runif(1,0.03,0.97),2)
z1 = round(qnorm(runif(1,0.03,0.97)),1)
u4 = round(runif(1,0.03,0.97),2)
z3 = round(qnorm(runif(1,0.03,0.97)),1)

u1 = round(pnorm(z1),3)
z2 = round(qnorm(u2),2)
u3 = round(pnorm(z3),3)
z4 = round(qnorm(u4),2)

p1 = c(0,u1)
p2 = c(0,u2)
p3 = c(u3,1)
p4 = c(u4,1)

#makepie(0,u1,"figs/one.pdf")
#makepie(0,u2,"figs/two.pdf")
#makepie(u3,1,"figs/three.pdf")
#makepie(u4,1,"figs/four.pdf")

u = c(u1,u2,u3,u4)
z = c(z1,z2,z3,z4)

qs = c(paste("Evaluate $P(Z <",z[1],")$",sep=""),
paste("Determine $z$ such that $P(Z < z) = ",u[2],"$",sep=""),
paste("Evaluate $P(Z >",z[3],")$",sep=""),
paste("Determine $z$ such that $P(Z > z) = ",1-u[4],"$",sep=""))

as = c(paste("$P(Z < ",z[1],") = \\fbox{",u[1],"}$"),
paste("$z = \\fbox{",z[2],"}$"),
paste("$P(Z > ",z[3],") = \\fbox{",1-u[3],"}$"),
paste("$z = \\fbox{",z[4],"}$"))

while(T){
  zs = round(qnorm(sort(runif(2,0.01,0.99))),1)
  us = round(pnorm(zs),3)
  P = round(us[2]-us[1],3)
  if(P>0.05){
    break
  }
}
qs = c(qs,paste("Evaluate $P(",zs[1],"<Z<",zs[2],")$"))
as = c(as,paste("$P(",zs[1],"<Z<",zs[2],") = \\fbox{",P,"}$"))
#makepie(us[1],us[2],"figs/five.pdf")
p5 = c(us[1],us[2])

z = round(qnorm(runif(1,0.54,0.96)),1)
qs = c(qs,paste("Evaluate $P(|Z|<",z,")$",sep=""))
as = c(as,paste("$P(|Z|<",z,") = \\fbox{",round(pnorm(z)*2-1,3),"}$",sep=""))
#makepie(pnorm(-z),pnorm(z),"figs/six.pdf")
p6 = c(pnorm(-z),pnorm(z))

u = round(runif(1,0.54,0.96),2)
P = round((u-0.5)*2,2)
z = round(qnorm(u),2)
qs = c(qs,paste("Determine $z$ such that $P(|Z|<z)=",P,"$",sep=""))
as = c(as,paste("$z = \\fbox{",z,"}$",sep=""))
p7 = c(pnorm(-z),pnorm(z))
#makepie(pnorm(-z),pnorm(z),"figs/seven.pdf")

z = round(qnorm(runif(1,0.54,0.96)),1)
qs = c(qs,paste("Evaluate $P(|Z|>",z,")$",sep=""))
as = c(as,paste("$P(|Z|>",z,") = \\fbox{",round(2*(pnorm(-z)),3),"}$",sep=""))
#makepie(pnorm(z),pnorm(-z),"figs/eight.pdf")
p8 = c(pnorm(z),pnorm(-z))

u = round(runif(1,0.54,0.96),2)
P = round((1-u)*2,2)
z = round(qnorm(u),2)
qs = c(qs,paste("Determine $z$ such that $P(|Z|>z)=",P,"$",sep=""))
as = c(as,paste("$z = \\fbox{",z,"}$",sep=""))
#makepie(pnorm(z),pnorm(-z),"figs/nine.pdf")
p9 = c(pnorm(z),pnorm(-z))

fignames = c("one","two","three","four","five","six","seven","eight","nine")

ord = 1:9

ppp = c(p1,p2,p3,p4,p5,p6,p7,p8,p9)

```

Question
========

The spinner below has two distributions. The outer distribution ($Z$) is the standard normal distribution. The inner distribution ($U$) is the standard uniform distribution, which also corresponds to the percentile. To answer the questions below, assume the spinner is equally likely to land in any direction.

```{r makespinner,echo=F,fig.width=6,fig.height=6}
library(latex2exp)
par(mar=c(0,0,0,0))
plot(1,1,type='n',axes=F,ann=F,xlim=c(-1,1),ylim=c(-1,1))
for(i in 1:100/100*2*pi){
  lines(c(cos(i)*0.75,cos(i)*0.8),c(sin(i)*0.75,sin(i)*0.8))
}
for(i in 1:10/10*2*pi){
  lines(c(cos(i)*0.65,cos(i)*0.8),c(sin(i)*0.65,sin(i)*0.8))
}
lines(c(0.5,0.8),c(0,0),lwd=3)
for(i in 1:9){
  text(cos(i/10*2*pi)*0.59,sin(i/10*2*pi)*0.59,i/10)
}
for(i in 0:100){
  if(i%%10 != 0){
    text(cos(i/100*2*pi)*0.69,sin(i/100*2*pi)*0.69,i/100,cex=0.6,srt=((i/100*360)+90)%%180-90)
  }
}
text(0.57,0.04,"0")
text(0.57,-0.04,"1")
for(zz in -30:30/10){
  lines(c(cos(pnorm(zz)*2*pi)*0.8,cos(pnorm(zz)*2*pi)*0.85),c(sin(pnorm(zz)*2*pi)*0.8,sin(pnorm(zz)*2*pi)*0.85))
}
for(zz in -20:20/10){
  if(zz%%1 != 0){
    text(cos(pnorm(zz)*2*pi)*0.89,sin(pnorm(zz)*2*pi)*0.89,zz,cex = 0.6)
  }
}
for(zz in -3:3){
  lines(c(cos(pnorm(zz)*2*pi)*0.8,cos(pnorm(zz)*2*pi)*0.95),c(sin(pnorm(zz)*2*pi)*0.8,sin(pnorm(zz)*2*pi)*0.95),lwd=2)
}
for(zz in -2:2){
    text(cos(pnorm(zz)*2*pi)*0.99,sin(pnorm(zz)*2*pi)*0.99,zz)
}
text(0.99,0.03,"-3")
text(0.99,-0.03,"3")
lines(cos(0:100/100*2*pi)*0.8,sin(0:100/100*2*pi)*0.8)

text(0.45,0,TeX("U"))
text(1.05,0,"Z")
```
\

Answerlist
--------
* `r qs[ord[1]]`
* `r qs[ord[2]]`
* `r qs[ord[3]]`
* `r qs[ord[4]]`
* `r qs[ord[5]]`
* `r qs[ord[6]]`
* `r qs[ord[7]]`
* `r qs[ord[8]]`
* `r qs[ord[9]]`


Solution
========

The following circles are meant to help visualize how to get the answers.

```{r a,echo=F,fig.height=5,fig.width=5,fig.cap=""}
par(mfrow=c(3,3))
for(i in 1:9){
  makepie(ppp[ord[i]*2-1],ppp[ord[i]*2])
  text(-0.95,0.95,c("a","b","c","d","e","f","g","h","i")[i])
}
```
\

Answerlist
-------
* `r as[ord[1]]`
* `r as[ord[2]]`
* `r as[ord[3]]`
* `r as[ord[4]]`
* `r as[ord[5]]`
* `r as[ord[6]]`
* `r as[ord[7]]`
* `r as[ord[8]]`
* `r as[ord[9]]`



Meta-information
============
extype: string
exsolution: yo
exname: UZ
extol: 0.01
