---
title: "Confidence Interval of Proportion"
header-includes:
   - \usepackage{arev}
output:
    pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# How to determine a confidence interval of a proportion

## Definitions

$$\begin{aligned}
p &= \text{population proportion (to be estimated)} \\
\hat{p} &= \text{sample proportion (actually measured)} \\
n &= \text{sample size} \\
\gamma &= \text{confidence level} = \text{chance that a confidence interval will include $p$} \\
\alpha &= \text{error rate} = 1-\gamma\\
\sigma_{\hat{p}} &= \text{standard error (standard deviation of sampling distribution)} \\
z^{\star} &= \text{standardized radius of interval} \\
ME &= \text{margin of error (radius of interval)} = z^{\star}\cdot\sigma_{\hat{p}} \\
\ell &= \text{percentile associated with $z^{\star}$} \\
LB &= \text{lower bound of confidence interval} = \hat{p}-ME \\
UB &= \text{upper bound of confidence interval} = \hat{p}+ME \\
CI &= \text{confidence interval} =[LB,\, UB]
\end{aligned}$$

## General Problem

- **Given:** $\hat{p}$, $n$, and $\gamma$ 
- **Find:** The lower and upper bounds of the confidence interval.

## General Procedure

We first determine $z^{\star}$ such that $P(|Z|<z^{\star})=\gamma$. The graphic below suggests the strategy: determine $\alpha$, find $\ell$, and use the $z$-table to get $z^{\star}$.

```{r visualize,fig.width=7,fig.height=3,fig.align='center'}
library(latex2exp)
par(mar=c(5,8,0,0))
plot(-300:300/100,dnorm(-300:300/100),axes=F,ann=F,type='l',ylim=c(0,0.5))
lines(c(-3,3),c(0,0))
z = 1.5
arrows(z,0.45,-3.2,0.45)
lines(c(z,z),c(0.45,0.43))
text((z-3)/2,0.47,TeX("percentile of $z^{*}$"))
lines(c(-z,-z),c(0,dnorm(-z)))
lines(c(z,z),c(0,dnorm(z)))
axis(1,c(-3:3,-1.5,1.5),c(-3:3,TeX("$-z^*$"),TeX("$z^*$")))
axis(1,line=3,c(0,-1.5,1.5),c(TeX("$\\hat{p}$"),TeX("$\\hat{p}-ME$"),TeX("$\\hat{p}+ME$")))
text(-1.9,0.03,TeX("$\\alpha/2$"))
text(1.9,0.03,TeX("$\\alpha/2$"))
text(0,0.1,TeX("$\\gamma$"),cex=2)
mtext("standardized scale:",1,1,adj=-.4)
mtext("proportion scale:",1,4,adj=-.1)
```


Determine error rate. This represents how often confidence intervals will miss the true population proportion. This error rate is a two-tail area.
$$\alpha = 1-\gamma $$
We can determine the percentile ($\ell$) of $z^{\star}$.
$$\ell = 1-\frac{\alpha}{2} $$
It should be mentioned that you could have gotten the percentile more directly.
$$\ell = \frac{\gamma+1}{2} $$
Use the $z$-table to get $z^*$.

We estimate the standard error. (Technically we should use $p$, not $\hat{p}$, but we only know $\hat{p}$. We assume $\hat{p}$ is close enough to $p$ for this estimation to be accurate.)
$$\sigma_{\hat{p}} = \sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$$
Determine the margin of error.
$$ME = z^{\star}\cdot \sigma_{\hat{p}}$$
Get the lower bound.
$$LB = \hat{p}-ME $$
Get the upper bound.
$$UB = \hat{p}+ME $$
Write the confidence interval in interval notation.
$$\fbox{$CI = [LB, UB] $}$$
We can summarize the procedure in two steps:

1. Determine $z^{\star}$ such that $P(|Z|<z^{\star})=\gamma$. It is helpful to know $\ell=\frac{\gamma+1}{2}$.
2. Use the following expression to find the bounds:
$$\hat{p}\pm z^{\star}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$$


