---
output:
  pdf_document: default
  html_document: default
---
```{r data generation, echo = FALSE, results = "hide"}
## DATA GENERATION
library(latex2exp)
options(scipen=999)
while(T){
  n = sample(5:8,1)
  x = signif(runif(n,1,10),2)
  scale = 10^sample(-1:1,1)
  b = runif(1,-2,2)*scale
  a = runif(1,1,10)*scale
  y = signif(a+b*x+rnorm(n,0,runif(1,0,2)*scale),2)
  if(sum(y<0)==0){break}
}
s = '\\begin{tabular}{|c|c|}\\hline\n $x$ & $y$ \\\\ \\hline \n'
for(i in 1:length(x)){
  s = paste(s,x[i]," & ",y[i]," \\\\ \n",sep="")
}
s = paste(s,'\\hline \n \\end{tabular}')
```

Question
========

You have collected the following data:

`r s`

\vfill

Please plot the data and a corresponding regression line.

\vfill
  
\
```{r makegridF,fig.width=5,fig.height=5,echo=F}
par(mar=c(0,0,0,0))
plot(0,0,type="n",axes=F,ann=F,xlim=c(0,1),ylim=c(0,1))
for(i in seq(0,1,0.01)){
  lines(c(i,i),c(0,1),lwd=0.3)
  lines(c(0,1),c(i,i),lwd=0.3)
}
for(i in seq(0,1,0.1)){
  lines(c(i,i),c(0,1),lwd=0.7)
  lines(c(0,1),c(i,i),lwd=0.7)
}
```
\

\vfill

Solution
========

Remember the formula for the correlation coefficient.

$$r = \frac{\sum x_i y_i - n\bar{x}\bar{y}}{(n-1)s_x s_y}$$

We calculate the necessary values.

```{r cakc, echo=F}
s2 = '\\begin{tabular}{|c|c|c|}\\hline\n $x$ & $y$ & $xy$ \\\\ \\hline \n'
for(i in 1:length(x)){
  s2 = paste(s2,x[i]," & ",y[i]," & ",x[i]*y[i]," \\\\ \n",sep="")
}
s2 = paste(s2,'\\hline \\hline \n $\\sum x = ',sum(x),'$ & $\\sum y = ',sum(y),'$ & $\\sum x_i y_i = ',sum(x*y),'$ \\\\ \\hline \n',sep="")
s2 = paste(s2,'\\hline \n $\\bar{x} = ',signif(mean(x),3),'$ & $\\bar{ y} = ',signif(mean(y),3),'$ &  \\\\ \\hline \n',sep="")
s2 = paste(s2,'\\hline \n $s_x = ',signif(sd(x),3),'$ & $s_y = ',signif(sd(y),3),'$ &  \\\\ \\hline \n',sep="")
s2 = paste(s2,' \n \\end{tabular}')

sx = signif(sd(x),3)
sy = signif(sd(y),3)
r = signif(cor(x,y),3)
b = signif(r*sy/sx,3)
a = signif(mean(y)-b*mean(x),3)
```

`r s2`

The regression line has the form
$$y = a + bx $$
So, $a$ is the $y$-intercept and $b$ is the slope. We have formulas to determine them:
$$b = r \frac{s_y}{s_x} =  `r r` \cdot \frac{`r sy`}{`r sx`} = `r b`$$
$$a = \bar{y}-b\bar{x} = `r signif(mean(y),3)`-`r b`\cdot`r signif(mean(x),3)` = `r a` $$ 
Our regression line:
$$y = `r a`+`r b`x $$

Make a plot.
\
```{r makeansF,fig.width=5,fig.height=5,echo=F}
possmax = c(0.1,0.2,0.5,1,2,5,10,20,50,100,200,500,1000,2000,5000,10000)
myl = min(possmax[possmax>max(max(y),a)])
plot(x,y,xlim=c(0,1)*10,ylim=c(0,myl))
for(i in seq(0,1,0.01)){
  lines(c(i,i)*10,c(0,1)*myl,lwd=0.3)
  lines(c(0,1)*10,c(i,i)*myl,lwd=0.3)
}
for(i in seq(0,1,0.1)){
  lines(c(i,i)*10,c(0,1)*myl,lwd=0.7)
  lines(c(0,1)*10,c(i,i)*myl,lwd=0.7)
}
abline(a,b)
```
\



Meta-information
============
extype: num
exsolution: `r x`
exname: binomial exact
extol: 0.01
