```{r data generation, echo = FALSE, results = "hide"}
types = c("skew left","skew right","bell","uniform","bimodal")
b1 = c(2, 0.5, 2, 1, 0.5)
b2 = c(0.5, 2, 2, 1, 0.5)
ord = sample(1:5)

names = c("F","G","H","I","J")
names = c("A","B","C","D","E")

ans1 = c("lower than","higher than","equal to","equal to","equal to")
cau = c("","","not","not","not")
```

Question
========

For **each** of the histograms:

* Determine if the mean is higher than, lower than, or equal to the median.
* Would you caution against using the mean?

```{r hists, echo=F}
par(mfrow=c(2,3))
for(i in 1:5){
  hist(rbeta(10000,b1[ord[i]],b2[ord[i]]),axes=F,xlab="measurement",ylab="density",main=names[i])
}
```

Answerlist
--------
* Answer both questions about distribution A.
* Answer both questions about distribution B.
* Answer both questions about distribution C.
* Answer both questions about distribution D.
* Answer both questions about distribution E.


Solution
========

Answerlist
--------
* The mean is `r ans1[ord[1]]` the median. Also, I would `r cau[ord[1]]` caution against using the mean.
* The mean is `r ans1[ord[2]]` the median. Also, I would `r cau[ord[2]]` caution against using the mean.
* The mean is `r ans1[ord[3]]` the median. Also, I would `r cau[ord[3]]` caution against using the mean.
* The mean is `r ans1[ord[4]]` the median. Also, I would `r cau[ord[4]]` caution against using the mean.
* The mean is `r ans1[ord[5]]` the median. Also, I would `r cau[ord[5]]` caution against using the mean.


Meta-information
============
extype: num
exsolution: 10
exname: calc s