---
title: "Statistics Overview"
output:
  ioslides_presentation:
    mathjax: local
    self_contained: false
---

## Outline

> - Randomness
> - Summarizing results
> - Probability (predicting results)
> - Central Limit Theorem
> - Inference (interpreting results)
> - Study Design

## Randomness

<video width="600" height="300" poster="videos/ryg_spins/ryg_spins_im.png" autoplay>
  <source src="videos/ryg_spins/ryg_spins.webm" type="video/webm">
</video>

> - Common random-number generators: dice, spinners, cards, coins, tops...
> - How long will it take to get home?
> - Some deterministic systems are treated as random.

## Randomness 2

We will study three types of generators: categorical, discrete, and continuous.




## Summarizing 

- Raw data is overwhelming.

```{r raw, fig.height=3,fig.width=3,echo=FALSE}
circ = function(r=.5,cen=0.5,n=1000,color=rgb(0,0,0,1)){
  angles = (1:n)/n*2*pi
  xs = cos(angles)*r+cen[1]
  ys = sin(angles)*r+cen[2]
  polygon(xs,ys,col = color)
}
par(mar = c(0,0,0,0))
plot(1, 1, type="n", axes=FALSE, ann=FALSE, xlim=c(0,1), ylim = c(0,1))
fig.align="center"
set.seed(123)
phits = runif(64)
pseps = c(0,1/3,25/60)
cols = c(rgb(0.3,1,0.3),rgb(1,1,0.3),rgb(1,0.3,0.3))
frequencies = c(0,0,0)
for(ii in 1:length(phits)){
  xx = 0.1+0.1*((ii-1)%%8)
  yy = 0.9-0.1*((ii-1)%/%8)
  res = sum(pseps <= phits[ii])
  frequencies[res] = frequencies[res]+1
  circ(r=0.1/2.2,cen=c(xx,yy),n=100,color=cols[res])
}
```


## Summarizing 2
- So, instead we summarize.

| Color | Frequency | Relative Frequency |
|:-----:|:---------:|:------------------:|
| Green | 23 |  $\frac{23}{64}\approx`r round(23/64,4)`$|
| Yellow| 6  | $\frac{6}{64}\approx`r round(6/64,4)`$|
| Red   | 35 | $\frac{35}{64}\approx`r round(35/64,4)`$|

## Summarizing 3
- So, instead we summarize. We are visual creatures...
```{r bargraph, echo=FALSE}
barplot(frequencies, names.arg = c("green","yellow","red"), col = c(rgb(0.3,1,0.3),rgb(1,1,0.3),rgb(1,0.3,0.3)),cex.names = 1.5, ylab = "frequency")
```

## Probability
A wedge's size determines its long-run relative frequency.

<video width="800" height="400" autoplay controls>
  <source src="videos/running_rel_freq/running.webm" type="video/webm">
</video>

## Probability 2
Long-run average winnings can be calculated as a weighted average.



