---
output:
  pdf_document: default
  html_document: default
---

```{r data generation, echo = FALSE, results = "hide"}

u2 = round(runif(1,0.01,0.99),2)
z1 = round(qnorm(runif(1,0.01,0.99)),1)
u4 = round(runif(1,0.01,0.99),2)
z3 = round(qnorm(runif(1,0.01,0.99)),1)

u1 = round(pnorm(z1),3)
z2 = round(qnorm(u2),2)
u3 = round(pnorm(z3),3)
z4 = round(qnorm(u4),2)

u = c(u1,u2,u3,u4)
z = c(z1,z2,z3,z4)

qs = c(paste("Evaluate $P(Z <",z[1],")$.",sep=""),
paste("Determine $z$ such that $P(Z < z) = ",u[2],"$.",sep=""),
paste("Evaluate $P(Z >",z[3],")$.",sep=""),
paste("Determine $z$ such that $P(Z > z) = ",1-u[4],"$.",sep=""))

as = c(paste("$P(Z < ",z[1],") = \\fbox{",u[1],"}$"),
paste("$z = \\fbox{",z[2],"}$"),
paste("$P(Z > ",z[3],") = \\fbox{",1-u[3],"}$"),
paste("$z = \\fbox{",z[4],"}$"))
while(T){
  zs = round(qnorm(sort(runif(2,0.01,0.99))),1)
  us = round(pnorm(zs),3)
  P = round(us[2]-us[1],3)
  if(P>0.02){
    break
  }
}
qs = c(qs,paste("Evaluate $P(",zs[1],"<Z<",zs[2],")$."))
as = c(as,paste("$P(",zs[1],"<Z<",zs[2],") = \\fbox{",P,"}$"))
ord = sample(1:5)

```

Question
========

The spinner below has two distributions. The outer distribution ($Z$) is the standard normal distribution. The inner distribution ($U$) is the standard uniform distribution, which also corresponds to the percentile. To answer the questions below, assume the spinner is equally likely to land in any direction.

```{r makespinner,echo=F,fig.width=6,fig.height=6}
library(latex2exp)
par(mar=c(0,0,0,0))
plot(1,1,type='n',axes=F,ann=F,xlim=c(-1,1),ylim=c(-1,1))
for(i in 1:100/100*2*pi){
  lines(c(cos(i)*0.75,cos(i)*0.8),c(sin(i)*0.75,sin(i)*0.8))
}
for(i in 1:10/10*2*pi){
  lines(c(cos(i)*0.65,cos(i)*0.8),c(sin(i)*0.65,sin(i)*0.8))
}
lines(c(0.5,0.8),c(0,0),lwd=3)
for(i in 1:9){
  text(cos(i/10*2*pi)*0.59,sin(i/10*2*pi)*0.59,i/10)
}
for(i in 0:100){
  if(i%%10 != 0){
    text(cos(i/100*2*pi)*0.69,sin(i/100*2*pi)*0.69,i/100,cex=0.6,srt=((i/100*360)+90)%%180-90)
  }
}
text(0.57,0.04,"0")
text(0.57,-0.04,"1")
for(zz in -30:30/10){
  lines(c(cos(pnorm(zz)*2*pi)*0.8,cos(pnorm(zz)*2*pi)*0.85),c(sin(pnorm(zz)*2*pi)*0.8,sin(pnorm(zz)*2*pi)*0.85))
}
for(zz in -20:20/10){
  if(zz%%1 != 0){
    text(cos(pnorm(zz)*2*pi)*0.89,sin(pnorm(zz)*2*pi)*0.89,zz,cex = 0.6)
  }
}
for(zz in -3:3){
  lines(c(cos(pnorm(zz)*2*pi)*0.8,cos(pnorm(zz)*2*pi)*0.95),c(sin(pnorm(zz)*2*pi)*0.8,sin(pnorm(zz)*2*pi)*0.95),lwd=2)
}
for(zz in -2:2){
    text(cos(pnorm(zz)*2*pi)*0.99,sin(pnorm(zz)*2*pi)*0.99,zz)
}
text(0.99,0.03,"-3")
text(0.99,-0.03,"3")
lines(cos(0:100/100*2*pi)*0.8,sin(0:100/100*2*pi)*0.8)

text(0.45,0,TeX("U"))
text(1.05,0,"Z")
```
\

Answerlist
--------
* `r qs[ord[1]]`
* `r qs[ord[2]]`
* `r qs[ord[3]]`
* `r qs[ord[4]]`
* `r qs[ord[5]]`

Solution
========

Answerlist
-------
* `r as[ord[1]]`
* `r as[ord[2]]`
* `r as[ord[3]]`
* `r as[ord[4]]`
* `r as[ord[5]]`

Meta-information
============
extype: cloze
exclosetype: string|string|string|string|string
exsolution: `r as[ord[1]]` | `r as[ord[2]]` | `r as[ord[3]]` | `r as[ord[4]]` | `r as[ord[5]]`
exname: UZ
extol: 0.01
