---
output:
  pdf_document: default
  html_document: default
---
```{r data generation, echo = FALSE, results = "hide"}
zs = (-3*5):(3*5)/5
ys = 0:8*0.05
z = sample(1:(2*5)/5,1)
```

Question
========

The figure below shows the standard normal density. Each grid square represents 1% of probability.

```{r gridnorm,echo=F,fig.height=4,fig.width=10,fig.cap="",fig.path=paste("ug",sample(1:99999,1),sep="")}
plot(0,0,type="n",ann=F,axes=F,xlim=c(-3,3),ylim=c(0,0.4))
abline(h = ys,lty=2)
abline(v=zs,lty=2)
lines((-300:300)/100,dnorm((-300:300)/100))
axis(1,at=(-3):(3))
mtext("Z",side=1,line = 3)
lines(c(-3,3),c(0,0))
```
\

Answerlist
--------
* Estimate $z$ such that $P(|Z|>z) = `r round(2-2*pnorm(z),2)`$ by shading and counting.
* Determine $z$ such that $P(|Z|>z) = `r round(2-2*pnorm(z),2)`$ by using the $z$-table.


Solution
========

Answerlist
--------

* The shaded regions are shown below. \newline
```{r gridnormsol,echo=F,fig.height=4,fig.width=10,fig.cap="",fig.path=paste("ug",sample(1:99999,1),sep="")}
plot(0,0,type="n",ann=F,axes=F,xlim=c(-3,3),ylim=c(0,0.4))
abline(h = ys,lty=2)
abline(v=zs,lty=2)
xs = (-300:300)/100
dn = dnorm(xs)
lines(xs,dn)
lines(c(-3,3),c(0,0))
polygon(c(xs[xs<=-z],-z,-3),c(dn[xs<=-z],0,0),col = rgb(0.4,0.4,0.4,0.4))
polygon(c(xs[xs>=z],3,z),c(dn[xs>=z],0,0),col = rgb(0.4,0.4,0.4,0.4))
axis(1,at=(-3):(3))
mtext("Z",side=1,line = 3)
``` 
\
\newline
When you have shaded `r round(2-2*pnorm(z),2)*100` squares, starting at both tails, you should end near $z=`r z`$. Really, you want to shade `r round(round(2-2*pnorm(z),2)*100/2,3)` squares starting from the left and also `r round(round(2-2*pnorm(z),2)*100/2,3)` squares starting from the right.
* Each tail has half the two-tail area. So each tail has an area of `r round(2-2*pnorm(z),2)/2`. We can find the $z$ score with this left area...
$$z_{\text{left tail}} = `r round(qnorm(round(2-2*pnorm(z),2)/2),2)` $$
But, we want the positive value (the right tail's $z$ boundary).
$$z=\fbox{`r round(-qnorm(round(2-2*pnorm(z),2)/2),2)`}$$



Meta-information
============
extype: string
exsolution: yo
exname: marbles
extol: 0.01
